---
title: "Testing surface functions"
output: html_notebook
---

# Testing improvements to surface generation functions.

```{r setup, include = F}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

Load packages and functions

```{r}
library(tidyverse)
library(sf)
library(raster)
library(terra)
library(tictoc)
library(here)

source("../../code/fri_surface.R")
source("../../code/season_surface.R")
source("../../code/sev_surface.R")
source("../../code/patch_surface.R")
```

## Testing speed improvements from transitioning to terra

```{r}
## read in shapefiles
yose <- read_sf("../../data/spatial/yosemite.shp")
hucs <- read_sf("../../data/spatial/yose_sheds.shp")
yose_fires <- read_sf("../../data/spatial/yose_mtbs.shp") 

## let's test with one huc10 (Headwaters Merced River)
hmr = filter(hucs, Name == "Headwaters Merced River")
htr = filter(hucs, Name == "Headwaters Tuolumne River")
str = filter(hucs, Name == "South Fork Tuolumne River")
```

```{r}
## Testing new fri code
tic() 
fri_surface(landscape = hmr, #shapefile of landsape of interest
            fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
            fire_years = "Year", # label of the fire year column
            start_year = 1984, # Year prior to start of dataset (for landsat: 1983)
            end_year = 2018, # Year after end of dataset (for MTBS: 2018)
            decay_rate = 0.5, # Importance decay rate of the "invisible mosaic", between [0,1)
            out_raster = here("data/scratch/fri_hmr.tif")) #path for writing output rasters
toc()

rold = rast(here("data", "scratch", "fri_1804000801_old.tif"))
r = rast(here("data", "scratch", "fri_1804000801.tif"))
r = rast(here('data', 'scratch', 'fri_hmr.tif'))

## new season code
tic()
season_surface(landscape = hmr, #shapefile of landsape of interest
               fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
               fire_years = "Year", # label of the fire year column in fires shp
               fire_day = "jday",# label of fire day column in fires shp
               decay_rate = 0.5, # Importance decay rate of the "invisible mosaic", between [0,1)
               out_raster = here("data/scratch/sea_hmr.tif")) #path for writing output rasters
toc()
rold = rast(here("data", "scratch", "sea_1804000801.tif"))
r = rast(here('data', 'scratch', 'sea_hmr.tif'))

## new sev code
tic()
sev_surface(landscape = hmr, #shapefile of landsape of interest
            fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
            severity_dir = here("data/spatial/yose_mtbs_sev"), # directory where fire rasters are held
            fire_years = "Year", # label of the fire year column
            decay_rate = 0.5, # Importance decay rate of the "invisible mosaic", between [0,1)
            out_raster = here("data/scratch/sev_hmr.tif")) #path for writing output rasters
toc()
rold = rast(here("data", "scratch", "sev_1804000801.tif"))
r = rast(here('data', 'scratch', 'sev_hmr.tif'))

## new patch code
## hmr takes a real long time (at least with the old code) so picking a simpler fire history
htr = filter(hucs, Name == "Headwaters Tuolumne River")
tic()
patch_surface(landscape = hmr, #shapefile of landsape of interest
              fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
              severity_dir = here("data/spatial/yose_mtbs_sev"), # directory where fire rasters are held
              fire_years = "Year", # label of the fire year column
              decay_rate = 0.5, # Importance decay rate of the "invisible mosaic", between [0,1)
              out_raster = here("data/scratch/pat_hmr.tif")) #path for writing output rasters
toc()
## old used m^2, new uses ha
rold = rast(here("data", "scratch", "pat_1804000901.tif"))
r = rast(here('data', 'scratch', 'pat_htr.tif'))
r = rast(here('data', 'scratch', 'pat_hmr.tif'))
```

original fri surface function takes 14.75 seconds for the hmr watershed\
with terra run time is only 5.29 seconds. minor differences at the pixel level, likely do to how rasterizing is done?

Original sea surface function takes 14.21 seconds for the hmr watershed Updated version takes 3.93 seconds

Original sev surface function takes 20.07 seconds for the hmr watershed Updated version takes 3.4 seconds

Original patch surface function takes 43.42 seconds for the htr watershed (hmr was taking \> 1hr) Updated version takes 4.3 seconds!

## Adding a full (1) decay option to surface models  

```{r}
## read in shapefiles
yose <- read_sf("../../data/spatial/yosemite.shp")
hucs <- read_sf("../../data/spatial/yose_sheds.shp")
yose_fires <- read_sf("../../data/spatial/yose_mtbs.shp") 

## let's test with one huc10 (Headwaters Merced River)
hmr = filter(hucs, Name == "Headwaters Merced River")

## Testing new fri code; 11.06 sec
tic() 
fri_surface(landscape = hmr, #shapefile of landsape of interest
            fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
            fire_years = "Year", # label of the fire year column
            start_year = 1984, # Year prior to start of dataset (for landsat: 1983)
            end_year = 2018, # Year after end of dataset (for MTBS: 2018)
            decay_rate = 0.5, # Importance decay rate of the "invisible mosaic", between [0,1)
            out_raster = here("data/scratch/fri_hmr.tif")) #path for writing output rasters
toc()

r <- rast(here("data/scratch/fri_hmr.tif"))

## now try with a total decay; 8.5 sec
tic() 
fri_surface(landscape = hmr, #shapefile of landsape of interest
            fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
            fire_years = "Year", # label of the fire year column
            start_year = 1984, # Year prior to start of dataset (for landsat: 1983)
            end_year = 2018, # Year after end of dataset (for MTBS: 2018)
            decay_rate = 1, # Importance decay rate of the "invisible mosaic", between [0,1)
            out_raster = here("data/scratch/fri_hmr_decay1.tif")) #path for writing output rasters
toc()

r1 <- rast(here("data/scratch/fri_hmr_decay1.tif"))

## and now dropping the first interval
tic() 
fri_surface(landscape = hmr, #shapefile of landsape of interest
            fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
            fire_years = "Year", # label of the fire year column
            drop_censored_int = T,
            start_year = 1984, # Year prior to start of dataset (for landsat: 1983)
            end_year = 2018, # Year after end of dataset (for MTBS: 2018)
            decay_rate = 1, # Importance decay rate of the "invisible mosaic", between [0,1)
            out_raster = here("data/scratch/fri_hmr_decay1_drop1.tif")) #path for writing output rasters
toc()

r1.drop <- rast(here("data/scratch/fri_hmr_decay1_drop1.tif"))
```

OK, now doing the same for severity

```{r}
## new sev code
tic()
r <- sev_surface(
  landscape = hmr, #shapefile of landsape of interest
  fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
  severity_dir = here("data/spatial/yose_mtbs_sev"), # directory where fire rasters are held
  fire_years = "Year", # label of the fire year column
  decay_rate = 0.5 # Importance decay rate of the "invisible mosaic", between [0,1))
            ) 
toc()

## with a full decay (last fire only)
tic()
r1 <- sev_surface(
  landscape = hmr, #shapefile of landsape of interest
  fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
  severity_dir = here("data/spatial/yose_mtbs_sev"), # directory where fire rasters are held
  fire_years = "Year", # label of the fire year column
  decay_rate = 1 # Importance decay rate of the "invisible mosaic", between [0,1))
            ) 
toc()

## different watershed
tic()
r2 <- sev_surface(
  landscape = str, #shapefile of landsape of interest
  fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
  severity_dir = here("data/spatial/yose_mtbs_sev"), # directory where fire rasters are held
  fire_years = "Year", # label of the fire year column
  decay_rate = 0.5 # Importance decay rate of the "invisible mosaic", between [0,1))
            ) 
toc()

## old patch code
## hmr takes a real long time (at least with the old code) so picking a simpler fire history
htr = filter(hucs, Name == "Headwaters Tuolumne River")
rp <- rast(here("data/scratch/pat_hmr.tif"))

tic()
r.5 <- patch_surface(landscape = hmr, #shapefile of landsape of interest
              fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
              severity_dir = here("data/spatial/yose_mtbs_sev"), # directory where fire rasters are held
              fire_years = "Year", # label of the fire year column
              decay_rate = 0.5) # Importance decay rate of the "invisible mosaic", between [0,1)
toc()

tic()
r1.0 <- patch_surface(landscape = str, #shapefile of landsape of interest
              fires = yose_fires, #shapefile containing fire IDs associated with severity rasters
              severity_dir = here("data/spatial/yose_mtbs_sev"), # directory where fire rasters are held
              fire_years = "Year", # label of the fire year column
              decay_rate = 1.0) # Importance decay rate of the "invisible mosaic", between [0,1)
toc()
```
