---
title: "FDis_FRic_compare"
author: "Zack Steel"
# date: "5/31/2020"
output: rmarkdown::github_document
---

```{r setup, include = F}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = "..")
```

Comparing functional dispersion and functional richness (aka Minimum convex hull) as measures of pyrodiversity.  

I'm going to stick with Yosemite watersheds as with the other demos in this repository.  


```{r}
library(tidyverse)
# library(sf)
# library(raster)
# library(stars)
# library(fasterize)
# library(lwgeom)
# library(FD)
library(colorspace)
library(knitr)
library(ggdist)
library(cowplot)
library(GGally)
```

```{r, eval = F}
## read in shapefiles
yose <- read_sf("data/spatial/yosemite.shp")
hucs <- read_sf("data/spatial/yose_sheds.shp")
yose_fires <- read_sf("data/spatial/yose_mtbs.shp") 

ggplot() +
  geom_sf(data = yose, fill = "darkgreen", color = NA, alpha = 0.3) +
  geom_sf(data = hucs, fill = NA) +
  geom_sf(data = yose_fires, fill = "grey30", color = NA, alpha = 0.4) +
  theme_void()

as.data.frame(yose_fires) %>% 
  dplyr::select(Fire_ID, Fire_Name, Year, jday) %>% 
  head()
```

Let's calculate functional diversity/pyrodiversity first using individual traits.  

To calculate convex hull for all four traits need to set the number of PCoA axes to allow, becuase "species" can be very high, the default can have many axes. m = 4 might work.

Could also calculate Simpson's diversity using Vegan's diversity() function with some additional code.

```{r, eval = F}
source("code/pyrodiv_calc.R")
library(tictoc)
mround <- function(x, base, digits){
      base*round(x/base, digits = digits)
    } 

## get rasters
tic()
d <- lapply(1:nrow(hucs), function(i) {
  huc <- hucs[i,]
  id <- as.data.frame(huc)[1,"HUC10"]
  
  ## bring in trait rasters and adjust precision   
  ## E.g. Frequency - 1 year increments; Patch size - 1 log HA increments; 
  ## Severity - 0.5 CBI increments; Seasonality - 0.1 increments.

  fri_r <- paste0("data/spatial/yose_fri/fri_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 0)
  sea_r <- paste0("data/spatial/yose_sea/sea_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 1)
  sev_r <- paste0("data/spatial/yose_sev/sev_", id, ".tif") %>% 
    raster() %>% 
    mround(base = 5, digits = 1)
  pat_r <- paste0("data/spatial/yose_pat/pat_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 0)
  
  ## Bring in mask raster
  flam_r <- paste0("data/spatial/yose_flam/flam_", id, ".tif") %>% 
    raster()
  
  ## Run the function for each individual trait
  fri_d <- pyrodiv_calc(traits = fri_r,
                        frich = T,
                        mask = flam_r) %>% 
    mutate(huc = id,
           trait = "fri")
  sea_d <- pyrodiv_calc(traits = sea_r,
                        frich = T,
                        mask = flam_r) %>% 
    mutate(huc = id,
           trait = "sea")
  sev_d <- pyrodiv_calc(traits = sev_r,
                        frich = T,
                        mask = flam_r) %>% 
    mutate(huc = id,
           trait = "sev")
  pat_d <- pyrodiv_calc(traits = pat_r,
                        frich = T,
                        mask = flam_r) %>% 
    mutate(huc = id,
           trait = "pat")
  all_d <- pyrodiv_calc(traits = c(pat_r, sea_r, sev_r, pat_r),
                        frich = T,
                        pca_axes = 4,
                        mask = flam_r) %>% 
    mutate(huc = id,
           trait = "all")
  
  ## put together and return
  bind_rows(fri_d, sea_d, sev_d, pat_d, all_d) %>% 
    return()
}) %>% 
  bind_rows()
toc()

write.csv(d, "data/yose_trait_div.csv", row.names = F)
```

Let's take a look at what these look like.  

```{r, eval = F}
## Single trait diversity
d <- read.csv("../pyrodiv/data/intermediate/huc10_covs/traits_expanded.csv")
## Multi-trait diversity - just 30m res
dm <- read.csv("../pyrodiv/data/intermediate/huc10_covs/pyrodiv_expanded.csv") %>% 
  filter(fct == 1) %>% 
  mutate(trait = "all")
## Bring in huc info including flammable area
hmeta <- read.csv("../pyrodiv/data/intermediate/huc10_covs/burn_area.csv") %>% 
  dplyr::select(id = huc10, Area = flam_ha)
## Scale everything for easier comparison
# d2 <- dplyr::select(d, huc, FRic, FEve, FDis, RaoQ, simpson, trait) %>% 
#   group_by(trait) %>% 
#   mutate(FRic = (FRic - mean(FRic, na.rm = T)) / sd(FRic, na.rm = T),
#          FEve = (FEve - mean(FEve, na.rm = T)) / sd(FEve, na.rm = T),
#          FDis = (FDis - mean(FDis, na.rm = T)) / sd(FDis, na.rm = T),
#          RaoQ = (RaoQ - mean(RaoQ, na.rm = T)) / sd(RaoQ, na.rm = T),
#          simpson = (simpson - mean(simpson, na.rm = T)) / sd(simpson, na.rm = T)) %>% 
#   pivot_longer(cols = FRic:simpson, values_to = "value", names_to = "metric")

# ggplot(d2, aes(x = metric, y = value, fill = metric)) +
#   facet_grid(~trait, scales = "free") +
#   geom_violin(na.rm = TRUE)

## Add in huc area and look at correlations
d3 <- bind_rows(d, dm) %>% 
  merge(hmeta, by = "id") %>% 
  dplyr::select(FRic, FDis, Simpson = simpson, Area, trait) %>% 
  ## Scale everything for easier comparison
  mutate(FRic_l = log(FRic + 0.01),
         FDis_l = log(FDis + 0.01),
         area_l = log(Area))

## FRic has a real long right-hand tail with multi-trait
p <- filter(d3, trait == "all") %>% 
  dplyr::select(FRic_l, FDis, Simpson, Area) %>% 
  ggpairs() +
  theme_bw() +
  # theme(text = element_text(size = 14)) +
  ggtitle("Four-trait index")
p2 <- filter(d3, trait == "fri") %>% 
  dplyr::select(FRic, FDis, Simpson, Area) %>% 
  ggpairs() +
  theme_bw() +
  ggtitle("Fire Frequency")
p3 <- filter(d3, trait == "sea") %>% 
  dplyr::select(FRic, FDis, Simpson, Area) %>% 
  ggpairs() +
  theme_bw() +
  ggtitle("Seasonality")
p4 <- filter(d3, trait == "pat") %>% 
  dplyr::select(FRic, FDis, Simpson, Area) %>% 
  ggpairs() +
  theme_bw() +
  ggtitle("Patch Size")
p5 <- filter(d3, trait == "sev") %>% 
  dplyr::select(FRic, FDis, Simpson, Area) %>% 
  ggpairs() +
  theme_bw() +
  ggtitle("Severity")

## How often do watersheds have the full range of severity?
filter(d, trait == "sev") %>% 
  mutate(FRic = round(FRic, 3),
         full = ifelse(FRic == 3, 1, 0)) %>% 
  summarise(prop_full = sum(full) / n())

save_plot("figures/cor_all.png", p,
          base_height = 4, base_width = 7)
save_plot("figures/cor_fri.png", p2,
          base_height = 4, base_width = 7)
save_plot("figures/cor_sea.png", p3,
          base_height = 4, base_width = 7)
save_plot("figures/cor_pat.png", p4,
          base_height = 4, base_width = 7)
save_plot("figures/cor_sev.png", p5,
          base_height = 4, base_width = 7)
```

Look at how the different diversity metrics handle different resolutions.  

```{r, eval = F}
source("code/pyrodiv_calc.R")
library(tictoc)
mround <- function(x, base, digits){
      base*round(x/base, digits = digits)
} 

## get rasters
tic()
d_res <- lapply(1:nrow(hucs), function(i) {
  huc <- hucs[i,]
  id <- as.data.frame(huc)[1,"HUC10"]
  print(id)
  
  ## Bring in mask raster; do first here
  flam_r <- paste0("data/spatial/yose_flam/flam_", id, ".tif") %>% 
    raster()
  
  ## bring in trait rasters and adjust precision   
  ## E.g. Frequency - 1 year increments; Patch size - 1 log HA increments; 
  ## Severity - 0.5 CBI increments; Seasonality - 0.1 increments.

  fri_r <- paste0("data/spatial/yose_fri/fri_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 0) 
  sea_r <- paste0("data/spatial/yose_sea/sea_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 1) 
  sev_r <- paste0("data/spatial/yose_sev/sev_", id, ".tif") %>% 
    raster() %>% 
    mround(base = 5, digits = 1) 
  pat_r <- paste0("data/spatial/yose_pat/pat_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 0) 
  
  ts <- stack(fri_r, sea_r, sev_r, pat_r)
  names(ts) <- c("fri", "sea", "sev", "pat")

    ## projection mask if needed
    if(!(identical(crs(ts), crs(flam_r)) & identical(extent(ts), extent(flam_r)))) {
      flam_r <- suppressWarnings(projectRaster(from = flam_r, crs = crs(ts), res = res(ts), 
                             method = "ngb", alignOnly = F)) %>% 
        ## necessary to resample to align extents
        resample(y = ts, method = "ngb")
    }
  ## run mask
  ts <- mask(ts, flam_r) 

  print(1)
  ## run at full resolution
  all1 <- pyrodiv_calc(traits = ts,
                        frich = T,
                        pca_axes = 4) %>% 
    mutate(fct = 1)
  
  ## run with different aggregations
  print(2)
  ts2 <- aggregate(ts, 2, mean)
  ts2$fri <- round(ts2$fri, 0)
  ts2$sea <- round(ts2$sea, 1)
  ts2$sev <- mround(ts2$sev, base = 5, digits = 1)
  ts2$pat <- round(ts2$pat, 0)
  
  #### Need to round everything again after aggregating
  
  all2 <- pyrodiv_calc(traits = ts2,
                        frich = T,
                        pca_axes = 4) %>% 
    mutate(fct = 2)
  
  print(4)
  ts4 <- aggregate(ts, 4, mean)
  ts4$fri <- round(ts4$fri, 0)
  ts4$sea <- round(ts4$sea, 1)
  ts4$sev <- mround(ts4$sev, base = 5, digits = 1)
  ts4$pat <- round(ts4$pat, 0)
  
  all4 <- pyrodiv_calc(traits = ts4,
                        frich = T,
                        pca_axes = 4) %>% 
    mutate(fct = 4)
  
  print(8)
  ts8 <- aggregate(ts, 8, mean)
  ts8$fri <- round(ts8$fri, 0)
  ts8$sea <- round(ts8$sea, 1)
  ts8$sev <- mround(ts8$sev, base = 5, digits = 1)
  ts8$pat <- round(ts8$pat, 0)
  
  all8 <- pyrodiv_calc(traits = ts8,
                        frich = T,
                        pca_axes = 4) %>% 
    mutate(fct = 8)
  
  print(16)
  ts16 <- aggregate(ts, 16, mean)
  ts16$fri <- round(ts16$fri, 0)
  ts16$sea <- round(ts16$sea, 1)
  ts16$sev <- mround(ts16$sev, base = 5, digits = 1)
  ts16$pat <- round(ts16$pat, 0)
  
  all16 <- pyrodiv_calc(traits = ts16,
                        frich = T,
                        pca_axes = 4) %>% 
    mutate(fct = 16)
  
  ## put together and return
  bind_rows(all1, all2, all4, all8, all16) %>% 
    mutate(id = id) %>% 
    return()
}) %>% 
  bind_rows()
toc()

write.csv(d_res, "data/yose_pd_vres.csv", row.names = F)


## set up aggregation
# res <- expand.grid(id = hucs$HUC10, agg = c(1:2)) %>% 
#   tibble() %>% 
#   
#   ## get raster paths
#   mutate(id = as.character(id),
#          flam = paste0("data/spatial/yose_flam/flam_", id, ".tif"),
#          fri = paste0("data/spatial/yose_flam/fri_", id, ".tif"),
#          sea = paste0("data/spatial/yose_flam/sea_", id, ".tif"),
#          pat = paste0("data/spatial/yose_flam/pat_", id, ".tif"),
#          sev = paste0("data/spatial/yose_flam/sev_", id, ".tif")) %>% 
#   nest(flam:sev) %>% 
#   mutate(rasters = purrr::map(data, ~sapply(.x, stack)))
#   ## aggregate rasters
#   mutate(flam_r = purrr::map2(flam, agg, ~aggregate(raster(.x), 
#                                                     fact = .y, fun = mean)),
#          fri_r = purrr::map2(fri, agg, ~aggregate(raster(.x),
#                                                     fact = .y, fun = mean)),
#          sea_r = purrr::map2(sea, agg, ~aggregate(raster(.x), 
#                                                     fact = .y, fun = mean)),
#          pat_r = purrr::map2(pat, agg, ~aggregate(raster(.x), 
#                                                     fact = .y, fun = mean)),
#          sev_r = purrr::map2(sev, agg, ~aggregate(raster(.x), 
#                                                     fact = .y, fun = mean)))
# 
# ## calculate pyrodiversity
# res2 <- nest(res)
```

Bring in data for all hucs  

```{r}
library(tidyverse)
library(cowplot)
library(tidybayes)
library(lme4)
library(broom.mixed)
library(colorspace)

d0 <- read.csv("../pyrodiv/data/intermediate/huc10_covs/pyrodiv_expanded.csv") %>% 
  ## Convert expansion factor to cell area
  mutate(grain = (fct * 30)^2 / 10000,
         grain_l = log(grain),
         ## Normalise FRic
         FRic_l = log(FRic + 0.001),
         FDis_s = (FDis - mean(FDis)) / sd(FDis),
         FRic_s = (FRic_l - mean(FRic_l)) / sd(FRic_l),
         simp_s = (simpson - mean(simpson)) / sd(simpson))

## Add base pyrodiversity back in
base <- filter(d0, fct == 1) %>% 
  dplyr::select(id, 
                b_dis = FDis_s,
                b_ric = FRic_s,
                b_sim = simp_s)

d <- merge(d0, base)

# ggplot(d, aes(x = fct, y = FDis_l, group = id)) +
#   geom_point() +
#   geom_line()
# 
# ggplot(d, aes(x = fct, y = FRic, group = id)) +
#   geom_point() +
#   geom_line()

# p <- dplyr::select(d, id, FRic_l, FDis_l, simpson, fct) %>% 
#   mutate(res = (fct * 30)^2 / 10000) %>% 
#   pivot_longer(cols = FRic_l:simpson, names_to = "metric", values_to = "value") %>% 
#   ggplot(aes(x = res, y = value, group = id)) +
#   facet_grid(row = vars(metric), scales = "free") +
#   geom_point(alpha = .1) +
#   geom_line(alpha = .1) +
#   scale_x_log10() +
#   xlab("Cell size (hectares)") + ylab("Pyrodiversity") +
#   theme_bw() 

p2 <- dplyr::select(d, id, FRic = FRic_s, FDis = FDis_s, Simpson = simp_s, grain) %>% 
  ## omit one outlier (aggregation anomoly in a small watershed), but keep for modeling below
  filter(FDis != max(FDis)) %>% 
  pivot_longer(cols = FRic:Simpson, names_to = "metric", values_to = "value") %>% 
  ggplot(aes(x = grain, y = value, group = id, color = metric)) +
  facet_grid(row = vars(metric), scales = "free") +
  geom_point(alpha = .1) +
  geom_line(alpha = .1) +
  scale_colour_discrete_qualitative(name = "Metric", guide = F) +
  scale_x_log10() +
  xlab("Cell size (hectares)") + ylab("Pyrodiversity") +
  theme_bw() 



## What does this look like statistically?
m_dis <- lmer(FDis_s ~ grain + grain:b_dis + (1|id), data = d)
m_ric <- lmer(FRic_s ~ grain + grain:b_ric + (1|id), data = d)
m_sim <- lmer(simp_s ~ grain + grain:b_sim + (1|id), data = d)

tb <- tibble(metric = c("FDis", "FRic", "Simpson"),
             models = list(m_dis, m_ric, m_sim)) %>% 
  mutate(summary = map(models, ~tidy(.x, conf.int = T, conf.level = 0.9))) %>% 
  unnest(summary) %>% 
  dplyr::select(metric, term, estimate, conf.low, conf.high) %>% 
  filter(term %in% c("grain", "grain:b_dis", "grain:b_ric", "grain:b_sim")) %>% 
  mutate(term = ifelse(term %in% c("grain:b_dis", "grain:b_ric", "grain:b_sim"), 
                       "grain:pd", "grain"),
         term = fct_rev(term)) %>% 
  mutate_if(is.numeric, round, 4)

tb

dots <- ggplot(tb, aes(x = estimate, y = term, xmin = conf.low, xmax = conf.high,
                          color = metric)) +
  geom_pointintervalh(size = 6, position = position_dodge(width=0.3)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_colour_discrete_qualitative(name = "Metric") +
  theme_bw() +
  theme(legend.direction = "horizontal",
        legend.position = "top") +
  ylab(NULL)

save_plot(filename = "figures/vary_scales.png", p2,
          base_width = 6, base_height = 6, units = "in")
save_plot(filename = "figures/vary_scales_dots.png", dots,
          base_width = 4, base_height = 3, units = "in")
```


Look at how measures vary depending on species definition/cut-offs. Here just varying how severity is defined and using the Upper Merced.  

Having problems with varying the cutoffs...


```{r}
## get rasters
tic()

dd <- lapply(1:nrow(hucs), function(i) {
  huc <- hucs[i,]
  id <- as.data.frame(huc)[1,"HUC10"]
  
  ## bring in trait rasters and adjust precision   

  # fri_r <- paste0("data/spatial/yose_fri/fri_", id, ".tif") %>%
  #   raster() %>%
  #   round(digits = 0)
  # sea_r <- paste0("data/spatial/yose_sea/sea_", id, ".tif") %>%
  #   raster() %>%
  #   round(digits = 1)
  # pat_r <- paste0("data/spatial/yose_pat/pat_", id, ".tif") %>%
  #   raster() %>%
  #   round(digits = 0)
  
  ## Bring in mask raster
  flam_r <- paste0("data/spatial/yose_flam/flam_", id, ".tif") %>% 
    raster()
  
    ## Cut-offs = ordered classes when unburned is an additional class. Run for 4 through 10 classes
  sev_r <- paste0("data/spatial/yose_sev/sev_", id, ".tif") %>% 
    raster()
  
  sev_l <- tibble(huc = id, classes = seq(4, 10)) %>% 
    mutate(raster = purrr::map(classes, ~{sev_r %>% 
    cut(breaks = seq(0,3, length.out = .x),
        include.lowest = T)
    }),
    ## put back on cbi scale (0-3)
    raster2 = purrr::map(raster, ~(.x-1) / cellStats((.x-1), max) * cellStats(sev_r, max)),
    ## Calculate pyrodiersity metrics
    sev_d = purrr::map(raster2, ~pyrodiv_calc(traits = .x,
                                             frich = F,
                                             mask = flam_r) %>% 
                         mutate(trait = "sev")))
  
  
  ## pull out the relevant data
  sev_d <- unnest(sev_l, cols = sev_d) %>% 
    dplyr::select(-raster, -raster2)

  
  # all_d <- pyrodiv_calc(traits = c(pat_r, sea_r, sev_r, pat_r),
  #                       frich = T,
  #                       pca_axes = 4,
  #                       mask = flam_r) %>% 
  #   mutate(huc = id,
  #          trait = "all") %>% 
  #   dplyr::select(-raster)
  # 
  # ## put together and return
  # bind_rows(fri_d, sea_d, sev_d, pat_d, all_d) %>% 
  #   return()
}) %>% 
  bind_rows()
toc()

filter(dd, trait == "sev") %>% 
  dplyr::select(classes, FEve, FDis, RaoQ, simpson) %>% 
  ggpairs()
filter(dd, huc == "1804000801") %>% 
  dplyr::select(classes, FEve, FDis, RaoQ, simpson) %>% 
  ggpairs()
```

