---
title: "FDis_FRic_compare"
author: "Zack Steel"
# date: "5/31/2020"
output: rmarkdown::github_document
---

```{r setup, include = F}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = "..")
```

Comparing functional dispersion and functional richness (aka Minimum convex hull) as measures of pyrodiversity.  

I'm going to stick with Yosemite watersheds as with the other demos in this repository.  


```{r}
library(tidyverse)
library(sf)
library(raster)
library(stars)
library(fasterize)
library(lwgeom)
library(FD)
library(colorspace)
library(knitr)
library(ggdist)

## read in shapefiles
yose <- read_sf("data/spatial/yosemite.shp")
hucs <- read_sf("data/spatial/yose_sheds.shp")
yose_fires <- read_sf("data/spatial/yose_mtbs.shp") 

ggplot() +
  geom_sf(data = yose, fill = "darkgreen", color = NA, alpha = 0.3) +
  geom_sf(data = hucs, fill = NA) +
  geom_sf(data = yose_fires, fill = "grey30", color = NA, alpha = 0.4) +
  theme_void()

as.data.frame(yose_fires) %>% 
  dplyr::select(Fire_ID, Fire_Name, Year, jday) %>% 
  head()
```

Let's calculate functional diversity/pyrodiversity first using individual traits.  

I get an error when trying to calculate convex hull for all four traits. maybe too much dimensionality? May need to set the number of PCoA axes to allow, becuase "species" can be very high, the default can have many axes. m = 4 might work.

Could also calculate Simpson's diversity using Vegan's diversity() function with some additional code.

```{r, eval = F}
source("code/pyrodiv_calc.R")
library(tictoc)
mround <- function(x, base, digits){
      base*round(x/base, digits = digits)
    } 

## get rasters
tic()
d <- lapply(1:nrow(hucs), function(i) {
  huc <- hucs[i,]
  id <- as.data.frame(huc)[1,"HUC10"]
  
  ## bring in trait rasters and adjust precision   
  ## E.g. Frequency - 1 year increments; Patch size - 1 log HA increments; 
  ## Severity - 0.5 CBI increments; Seasonality - 0.1 increments.

  fri_r <- paste0("data/spatial/yose_fri/fri_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 0)
  sea_r <- paste0("data/spatial/yose_sea/sea_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 1)
  sev_r <- paste0("data/spatial/yose_sev/sev_", id, ".tif") %>% 
    raster() %>% 
    mround(base = 5, digits = 1)
  pat_r <- paste0("data/spatial/yose_pat/pat_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 0)
  
  ## Bring in mask raster
  flam_r <- paste0("data/spatial/yose_flam/flam_", id, ".tif") %>% 
    raster()
  
  ## Run the function for each individual trait
  fri_d <- pyrodiv_calc(traits = fri_r,
                        frich = T,
                        mask = flam_r) %>% 
    mutate(huc = id,
           trait = "fri")
  sea_d <- pyrodiv_calc(traits = sea_r,
                        frich = T,
                        mask = flam_r) %>% 
    mutate(huc = id,
           trait = "sea")
  sev_d <- pyrodiv_calc(traits = sev_r,
                        frich = T,
                        mask = flam_r) %>% 
    mutate(huc = id,
           trait = "sev")
  pat_d <- pyrodiv_calc(traits = pat_r,
                        frich = T,
                        mask = flam_r) %>% 
    mutate(huc = id,
           trait = "pat")
  all_d <- pyrodiv_calc(traits = c(pat_r, sea_r, sev_r, pat_r),
                        frich = T,
                        pca_axes = 4,
                        mask = flam_r) %>% 
    mutate(huc = id,
           trait = "all")
  
  ## put together and return
  bind_rows(fri_d, sea_d, sev_d, pat_d, all_d) %>% 
    return()
}) %>% 
  bind_rows()
toc()

write.csv(d, "data/yose_trait_div.csv", row.names = F)
```

Let's take a look at what these look like.  

```{r}
library(GGally)
d <- read.csv("data/yose_trait_div.csv")
## Scale everything for easier comparison
d2 <- dplyr::select(d, huc, FRic, FEve, FDis, RaoQ, simpson, trait) %>% 
  group_by(trait) %>% 
  mutate(FRic = (FRic - mean(FRic, na.rm = T)) / sd(FRic, na.rm = T),
         FEve = (FEve - mean(FEve, na.rm = T)) / sd(FEve, na.rm = T),
         FDis = (FDis - mean(FDis, na.rm = T)) / sd(FDis, na.rm = T),
         RaoQ = (RaoQ - mean(RaoQ, na.rm = T)) / sd(RaoQ, na.rm = T),
         simpson = (simpson - mean(simpson, na.rm = T)) / sd(simpson, na.rm = T)) %>% 
  pivot_longer(cols = FRic:simpson, values_to = "value", names_to = "metric")

ggplot(d2, aes(x = metric, y = value, fill = metric)) +
  facet_grid(~trait, scales = "free") +
  geom_violin(na.rm = TRUE)

## Add in huc area and look at correlations
d3 <- dplyr::select(hucs, huc = HUC10, sqkm = AreaSqKm) %>% 
  merge(d) %>% 
  st_drop_geometry()

p <- filter(d3, trait == "all") %>% 
  dplyr::select(FRic, FDis, simpson, sqkm) %>% 
  ggpairs() +
  ggtitle("Four-trait index")
p2 <- filter(d3, trait == "fri") %>% 
  dplyr::select(FRic, FDis, simpson, sqkm) %>% 
  ggpairs() +
  ggtitle("Fire Frequency")
p3 <- filter(d3, trait == "sea") %>% 
  dplyr::select(FRic, FDis, simpson, sqkm) %>% 
  ggpairs() +
  ggtitle("Seasonality")
p4 <- filter(d3, trait == "pat") %>% 
  dplyr::select(FRic, FDis, simpson, sqkm) %>% 
  ggpairs() +
  ggtitle("Patch Size")
p5 <- filter(d3, trait == "sev") %>% 
  dplyr::select(FRic, FDis, simpson, sqkm) %>% 
  ggpairs() +
  ggtitle("Severity")

save_plot("figures/core_all.png", p,
          base_height = 6, base_width = 8)
save_plot("figures/core_fri.png", p2,
          base_height = 6, base_width = 8)
save_plot("figures/core_sea.png", p3,
          base_height = 6, base_width = 8)
save_plot("figures/core_pat.png", p4,
          base_height = 6, base_width = 8)
save_plot("figures/core_sev.png", p5,
          base_height = 6, base_width = 8)
```

Look at how the different diversity metrics handle different resolutions.  

```{r}
source("code/pyrodiv_calc.R")
library(tictoc)
mround <- function(x, base, digits){
      base*round(x/base, digits = digits)
} 

## get rasters
tic()
d_res <- lapply(1:nrow(hucs), function(i) {
  huc <- hucs[i,]
  id <- as.data.frame(huc)[1,"HUC10"]
  print(id)
  
  ## Bring in mask raster; do first here
  flam_r <- paste0("data/spatial/yose_flam/flam_", id, ".tif") %>% 
    raster()
  
  ## bring in trait rasters and adjust precision   
  ## E.g. Frequency - 1 year increments; Patch size - 1 log HA increments; 
  ## Severity - 0.5 CBI increments; Seasonality - 0.1 increments.

  fri_r <- paste0("data/spatial/yose_fri/fri_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 0) 
  sea_r <- paste0("data/spatial/yose_sea/sea_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 1) 
  sev_r <- paste0("data/spatial/yose_sev/sev_", id, ".tif") %>% 
    raster() %>% 
    mround(base = 5, digits = 1) 
  pat_r <- paste0("data/spatial/yose_pat/pat_", id, ".tif") %>% 
    raster() %>% 
    round(digits = 0) 
  
  ts <- stack(fri_r, sea_r, sev_r, pat_r)
  names(ts) <- c("fri", "sea", "sev", "pat")

    ## projection mask if needed
    if(!(identical(crs(ts), crs(flam_r)) & identical(extent(ts), extent(flam_r)))) {
      flam_r <- suppressWarnings(projectRaster(from = flam_r, crs = crs(ts), res = res(ts), 
                             method = "ngb", alignOnly = F)) %>% 
        ## necessary to resample to align extents
        resample(y = ts, method = "ngb")
    }
  ## run mask
  ts <- mask(ts, flam_r) 

  print(1)
  ## run at full resolution
  all1 <- pyrodiv_calc(traits = ts,
                        frich = T,
                        pca_axes = 4) %>% 
    mutate(fct = 1)
  
  ## run with different aggregations
  print(2)
  ts2 <- aggregate(ts, 2, mean)
  ts2$fri <- round(ts2$fri, 0)
  ts2$sea <- round(ts2$sea, 1)
  ts2$sev <- mround(ts2$sev, base = 5, digits = 1)
  ts2$pat <- round(ts2$pat, 0)
  
  #### Need to round everything again after aggregating
  
  all2 <- pyrodiv_calc(traits = ts2,
                        frich = T,
                        pca_axes = 4) %>% 
    mutate(fct = 2)
  
  print(4)
  ts4 <- aggregate(ts, 4, mean)
  ts4$fri <- round(ts4$fri, 0)
  ts4$sea <- round(ts4$sea, 1)
  ts4$sev <- mround(ts4$sev, base = 5, digits = 1)
  ts4$pat <- round(ts4$pat, 0)
  
  all4 <- pyrodiv_calc(traits = ts4,
                        frich = T,
                        pca_axes = 4) %>% 
    mutate(fct = 4)
  
  print(8)
  ts8 <- aggregate(ts, 8, mean)
  ts8$fri <- round(ts8$fri, 0)
  ts8$sea <- round(ts8$sea, 1)
  ts8$sev <- mround(ts8$sev, base = 5, digits = 1)
  ts8$pat <- round(ts8$pat, 0)
  
  all8 <- pyrodiv_calc(traits = ts8,
                        frich = T,
                        pca_axes = 4) %>% 
    mutate(fct = 8)
  
  print(16)
  ts16 <- aggregate(ts, 16, mean)
  ts16$fri <- round(ts16$fri, 0)
  ts16$sea <- round(ts16$sea, 1)
  ts16$sev <- mround(ts16$sev, base = 5, digits = 1)
  ts16$pat <- round(ts16$pat, 0)
  
  all16 <- pyrodiv_calc(traits = ts16,
                        frich = T,
                        pca_axes = 4) %>% 
    mutate(fct = 16)
  
  ## put together and return
  bind_rows(all1, all2, all4, all8, all16) %>% 
    mutate(id = id) %>% 
    return()
}) %>% 
  bind_rows()
toc()

write.csv(d_res, "data/yose_pd_vres.csv", row.names = F)


## set up aggregation
# res <- expand.grid(id = hucs$HUC10, agg = c(1:2)) %>% 
#   tibble() %>% 
#   
#   ## get raster paths
#   mutate(id = as.character(id),
#          flam = paste0("data/spatial/yose_flam/flam_", id, ".tif"),
#          fri = paste0("data/spatial/yose_flam/fri_", id, ".tif"),
#          sea = paste0("data/spatial/yose_flam/sea_", id, ".tif"),
#          pat = paste0("data/spatial/yose_flam/pat_", id, ".tif"),
#          sev = paste0("data/spatial/yose_flam/sev_", id, ".tif")) %>% 
#   nest(flam:sev) %>% 
#   mutate(rasters = purrr::map(data, ~sapply(.x, stack)))
#   ## aggregate rasters
#   mutate(flam_r = purrr::map2(flam, agg, ~aggregate(raster(.x), 
#                                                     fact = .y, fun = mean)),
#          fri_r = purrr::map2(fri, agg, ~aggregate(raster(.x),
#                                                     fact = .y, fun = mean)),
#          sea_r = purrr::map2(sea, agg, ~aggregate(raster(.x), 
#                                                     fact = .y, fun = mean)),
#          pat_r = purrr::map2(pat, agg, ~aggregate(raster(.x), 
#                                                     fact = .y, fun = mean)),
#          sev_r = purrr::map2(sev, agg, ~aggregate(raster(.x), 
#                                                     fact = .y, fun = mean)))
# 
# ## calculate pyrodiversity
# res2 <- nest(res)
```

```{r}
library(cowplot)
ggplot(d_res, aes(x = fct, y = FDis, group = id)) +
  geom_point() +
  geom_line()

ggplot(d_res, aes(x = fct, y = FRic, group = id)) +
  geom_point() +
  geom_line()

p <- dplyr::select(d_res, id, FRic, FDis, simpson, fct) %>% 
  mutate(res = (fct * 30)^2 / 10000) %>% 
  pivot_longer(cols = FRic:simpson, names_to = "metric", values_to = "value") %>% 
  ggplot(aes(x = res, y = value, group = id)) +
  facet_grid(row = vars(metric), scales = "free") +
  geom_point() +
  geom_line() +
  scale_x_log10() +
  xlab("Cell size (hectares)") + ylab("Pyrodiversity") +
  theme_bw() 

save_plot(filename = "figures/vary_scales.png", p,
          base_width = 6, base_height = 6, units = "in")
```


Look at how measures vary depending on species definition/cut-offs. Here just varying how severity is defined and using the Upper Merced.  

Having problems with varying the cutoffs...


```{r}
## get rasters
tic()

dd <- lapply(1:nrow(hucs), function(i) {
  huc <- hucs[i,]
  id <- as.data.frame(huc)[1,"HUC10"]
  
  ## bring in trait rasters and adjust precision   

  # fri_r <- paste0("data/spatial/yose_fri/fri_", id, ".tif") %>%
  #   raster() %>%
  #   round(digits = 0)
  # sea_r <- paste0("data/spatial/yose_sea/sea_", id, ".tif") %>%
  #   raster() %>%
  #   round(digits = 1)
  # pat_r <- paste0("data/spatial/yose_pat/pat_", id, ".tif") %>%
  #   raster() %>%
  #   round(digits = 0)
  
  ## Bring in mask raster
  flam_r <- paste0("data/spatial/yose_flam/flam_", id, ".tif") %>% 
    raster()
  
    ## Cut-offs = ordered classes when unburned is an additional class. Run for 4 through 10 classes
  sev_r <- paste0("data/spatial/yose_sev/sev_", id, ".tif") %>% 
    raster()
  
  sev_l <- tibble(huc = id, classes = seq(4, 10)) %>% 
    mutate(raster = purrr::map(classes, ~{sev_r %>% 
    cut(breaks = seq(0,3, length.out = .x),
        include.lowest = T)
    }),
    ## put back on cbi scale (0-3)
    raster2 = purrr::map(raster, ~(.x-1) / cellStats((.x-1), max) * cellStats(sev_r, max)),
    ## Calculate pyrodiersity metrics
    sev_d = purrr::map(raster2, ~pyrodiv_calc(traits = .x,
                                             frich = F,
                                             mask = flam_r) %>% 
                         mutate(trait = "sev")))
  
  
  ## pull out the relevant data
  sev_d <- unnest(sev_l, cols = sev_d) %>% 
    dplyr::select(-raster, -raster2)

  
  # all_d <- pyrodiv_calc(traits = c(pat_r, sea_r, sev_r, pat_r),
  #                       frich = T,
  #                       pca_axes = 4,
  #                       mask = flam_r) %>% 
  #   mutate(huc = id,
  #          trait = "all") %>% 
  #   dplyr::select(-raster)
  # 
  # ## put together and return
  # bind_rows(fri_d, sea_d, sev_d, pat_d, all_d) %>% 
  #   return()
}) %>% 
  bind_rows()
toc()

filter(dd, trait == "sev") %>% 
  dplyr::select(classes, FEve, FDis, RaoQ, simpson) %>% 
  ggpairs()
filter(dd, huc == "1804000801") %>% 
  dplyr::select(classes, FEve, FDis, RaoQ, simpson) %>% 
  ggpairs()
```

